---
title: "Cheem sensitivity"
author: "Nicholas Spyrison"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
editor_options:
  chunk_output_type: console
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo       = TRUE, ## code
  include    = TRUE, ## plots
  eval       = TRUE, ## chunk
  message    = FALSE,
  warning    = FALSE,
  error      = FALSE,
  collapse   = TRUE,
  cache      = TRUE,
  cache.lazy = FALSE
)
## Set directory relative to package
knitr::opts_knit$set(root.dir = '..')
```

# Requirements

```{R}
## Primary work packages
require("DALEX")
require("treeshap")
require("caret") ## Encoding class vars
require("spinifex")
require("tourr")
require("ggplot2")
## EDA and utility
require("palmerpenguins") ## penguins dataset
require("GGally")
require("tictoc")
require("patchwork")
require("plotly")
## Local functions
source("./apps/cheem/trees_of_cheem.r")   ## Cheem functions
source("./apps/cheem/spinifex_ggproto.r") ## New (spinifex) ggproto_* api
## Not run, open local function files
if(F){
  file.edit("./apps/cheem/trees_of_cheem.r")
  file.edit("./apps/cheem/spinifex_ggproto.r")
}
```

# Data, palmerpenguins::penguins


## Data setup

```{r}
raw <- palmerpenguins::penguins ## 5 missings, visdat::vis_miss(raw)
raw_rmna <- raw[!is.na(raw$sex),]
xdat <- spinifex::scale_sd(raw_rmna[, 3:6])
clas1 <- raw_rmna$species
clas2 <- raw_rmna$sex
table(clas1) ## Unbalenced island obs
table(clas2) ## Balanced sex

## Encode classes {caret}
library("caret")
dv <- caret::dummyVars(" ~ .", data = raw_rmna)
encoded <- data.frame(predict(dv, raw_rmna))
str(encoded)
detach("package:caret", unload = TRUE)
encoded <- cbind(encoded[, 1:3], encoded[, 11:12], xdat)

## Additional setup not used in local attribution matrix
new_obs_x <- rnorm_observation(dat)
```

## EDA glimse

```{r}
bas_pca  <- basis_pca(xdat)
bas_olda <- basis_olda(datx, clas1)

require("patchwork")
view_frame(bas_pca, data = xdat, 
           aes_args = list(color = clas1, shape = clas2)) + 
  ggtitle("PC1:2") +
view_frame(bas_olda, data = xdat, 
           aes_args = list(color = clas1, shape = clas2)) + 
  ggtitle("OLDA1:2")

# bas_olda4 <- basis_olda(xdat, clas1, d = 4)
# proj <- xdat %*% bas_olda4
# colnames(proj) <- paste0("OLDA", 1:4)
# GGally::ggpairs(as.data.frame(proj), 
#                 mapping = aes(color = clas1),
#                 #, shap = clas2), ## Artifacts when using both classes... 
#                 lower = list(continuous = wrap("points", alpha = 0.3)))
```


## RF Model, shap df

```{r}
rf_test <- as.integer(clas1 == levels(clas1)[1])
.p <- ncol(xdat)
.rf <- randomForest::randomForest(rf_test~.,
                                  data = data.frame(rf_test, xdat),
                                  mtry = if(is.discrete(clas1)) sqrt(.p) else .p / 3L)

shap_df <- treeshap_df(.rf, xdat)
```

# Compare variable obs and SHAP obs

Keep in mind this is testing for the first level of species.

## Scatterplot matrix

```{r}
require("patchwork")
str(xdat)
str(shap_df)

g1 <- GGally::ggpairs(as.data.frame(xdat), mapping = aes(color = clas1),
                      lower = list(continuous = wrap("points", alpha = 0.3)))
g2 <- GGally::ggpairs(as.data.frame(shap_df), mapping = aes(color = clas1),
                      lower = list(continuous = wrap("points", alpha = 0.3)))
g1 + g2
```


## NMDS

Phrasing the question such as ~"explore the sensitivity of the SHAP values" reminds me of observational vacinity/distance, which reminds me of MDS, so let's look at NMDS with `MASS::isoMDS`

```{r}
require("spinifex")
require("MASS")
require('ggfortify')

dim(spinifex::PimaIndiansDiabetes_wide)
dat <- spinifex::PimaIndiansDiabetes_wide

nmds <- MASS::isoMDS(dist(dat)) ## see nmds$points especially
plot(nmds$points, type = "n")
text(nmds$points, labels = as.character(1:nrow(dat)))

#plotly::ggplotly(gg)
```


## Grand tour


```{r}


```

