---
title: "Cheem sensitivity"
author: "Nicholas Spyrison"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
editor_options:
  chunk_output_type: console
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo       = TRUE, ## code
  include    = TRUE, ## plots
  eval       = TRUE, ## chunk
  message    = FALSE,
  warning    = FALSE,
  error      = FALSE,
  collapse   = TRUE,
  cache      = TRUE,
  cache.lazy = FALSE
)
## Set directory relative to package
knitr::opts_knit$set(root.dir = '..')
```

# Requirements

```{R}
## Primary work packages
require("DALEX")
require("treeshap")
require("caret") ## Encoding class vars
require("spinifex")
require("tourr")
require("ggplot2")
## EDA and utility
require("palmerpenguins") ## penguins dataset
require("GGally")
require("tictoc")
require("patchwork")
require("plotly")
## Local functions
source("./apps/cheem/trees_of_cheem.r")   ## Cheem functions
source("./apps/cheem/spinifex_ggproto.r") ## New (spinifex) ggproto_* api
## Not run, open local function files
if(F){
  file.edit("./apps/cheem/trees_of_cheem.r")
  file.edit("./apps/cheem/spinifex_ggproto.r")
}

```

# Data, palmerpenguins::penguins


## Data setup

```{r}
raw <- palmerpenguins::penguins ## 5 missings, visdat::vis_miss(raw)
raw_rmna <- raw[!is.na(raw$sex),]
xdat <- spinifex::scale_sd(raw_rmna[, 3:6])
clas1 <- raw_rmna$species
clas2 <- raw_rmna$sex
table(clas1) ## Unbalanced island obs
table(clas2) ## Balanced sex

## Encode classes {caret}
library("caret")
dv <- caret::dummyVars(" ~ .", data = raw_rmna)
encoded <- data.frame(predict(dv, raw_rmna))
str(encoded)
detach("package:caret", unload = TRUE)
encoded <- cbind(encoded[, 1:3], encoded[, 11:12], xdat)

## Additional setup not used in local attribution matrix
new_obs_x <- rnorm_observation_of(xdat)
```

## Check data space

```{r}
bas_pca  <- basis_pca(xdat)
bas_olda <- basis_olda(xdat, clas1)

require("patchwork")
view_frame(bas_pca, data = xdat, 
           aes_args = list(color = clas1, shape = clas2)) + 
  ggtitle("PC1:2") +
view_frame(bas_olda, data = xdat, 
           aes_args = list(color = clas1, shape = clas2)) + 
  ggtitle("OLDA1:2")
```


## RF Model, shap df

```{r}
.p <- ncol(xdat)
.lvls <- levels(as.factor(clas1))
shap_df <- data.frame(matrix(NA, ncol = .p, nrow = nrow(xdat))) ## Init
#.mute <- sapply(1L:length(.lvls), function(i){
for(i in 1L:length(.lvls)){
  .bool_test <- clas1 == .lvls[1]
  .int_test <- as.integer(.bool_test)
  ## RF on full xdat
  .rf <- randomForest::randomForest(.int_test~.,
                                    data = data.frame(.int_test, xdat),
                                    mtry = if(is.discrete(clas1)) sqrt(.p) else .p / 3L)
  ## treeshap_df of only i-th class level obs.
  .in_class_obs <- xdat[.bool_test, ]
  .ret <- treeshap_df(.rf, .in_class_obs)
  shap_df[.bool_test, ] <- as.data.frame(.ret) ## must cast as data.frame.
}
#})
```

# Compare variable obs and SHAP obs

Keep in mind this is testing for the first level of species, Adelie.

## Scatterplot matrix

```{r}
str(xdat)
str(shap_df)

g1 <- GGally::ggpairs(as.data.frame(xdat), mapping = aes(color = clas1),
                      lower = list(continuous = wrap("points", alpha = 0.3))) +
  ggtitle("Variable space")
g2 <- GGally::ggpairs(as.data.frame(shap_df), mapping = aes(color = clas1),
                      lower = list(continuous = wrap("points", alpha = 0.3))) +
  ggtitle("SHAP-Adelie space")
g1 ## Can't patchwork a ggplotmatrix.
g2
```


## NMDS

Phrasing the question such as ~"explore the sensitivity of the SHAP values" reminds me of observational vicinity/distance, which reminds me of MDS, so let's look at NMDS with `MASS::isoMDS`

```{r}
require("MASS")
nmds_xdat <- MASS::isoMDS(dist(xdat))$points %>%
  as.data.frame() ## see nmds$points especially
nmds_shap <- MASS::isoMDS(dist(shap_df))$points %>%
   as.data.frame() ## see nmds$points especially
colnames(nmds_xdat) <- colnames(nmds_shap) <- paste0("NMDS", 1:2)
nmds_xdat$rownum <- 1:nrow(xdat)
nmds_shap$rownum <- 1:nrow(xdat)
g3 <- nmds_xdat %>%
  ggplot(aes(NMDS1, NMDS2, color = clas1, shap = clas2, rownum = rownum)) +
  geom_point() +
  theme_minimal() +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  ggtitle("NMDS of variable space")
g4 <- nmds_shap %>%
  ggplot(aes(NMDS1, NMDS2, color = clas1, shap = clas2, rownum = 1:nrow(xdat))) +
  geom_point() +
  theme_minimal() +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  ggtitle("NMDS of SHAP-Adelie space")
# require("patchwork")
# g3 + g4
plotly::ggplotly(g3)
plotly::ggplotly(g4)
```


## Grand tour


```{r}
xdat_scaled <- spinifex::scale_sd(xdat)
shap_scaled <- spinifex::scale_sd(shap_df)


gt_path <- save_history(
  xdat,
  tour_path = grand_tour(),
  max_bases = 10
)

ggt1 <- ggplot_tour(gt_path, xdat_scaled, angle = .1) +
  ggproto_basis_axes() +
  ggproto_data_background() +
  ggproto_data_points(aes_args = list(color = clas1, shape = clas2),
                      identity_args = list(size = 1.5, alpha = .7)) + 
  ggtitle("Variable space, sd scaled")
ggt2 <- ggplot_tour(gt_path, shap_scaled, angle = .1) +
  ggproto_basis_axes() +
  ggproto_data_background() +
  ggproto_data_points(aes_args = list(color = clas1, shape = clas2),
                      identity_args = list(size = 1.5, alpha = .7)) + 
  ggtitle("SHAP-Adelie space, sd scaled")

animate_plotly(ggt1)
animate_plotly(ggt2)
```

