---
title: "Trees of Cheem"
author: "Nicholas Spyrison"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
editor_options: 
  chunk_output_type: console
---
<!--   #rmarkdown::html_vignette -->
<!-- # vignette: > -->
<!-- #   %\VignetteIndexEntry{spinifex} -->
<!-- #   %\VignetteEngine{knitr::rmarkdown} -->
<!-- #   %\VignetteEncoding{UTF-8}{inputenc} -->
<!-- #   \usepackage[utf8]{inputenc} -->

<!-- #Example vignette: 
https://github.com/njtierney/naniar/blob/master/vignettes/getting-started-w-naniar.Rmd -->
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo       = TRUE, ## code
  include    = TRUE, ## plots
  eval       = TRUE, ## chunk
  message    = FALSE,
  warning    = FALSE,
  error      = FALSE,
  collapse   = TRUE,
  cache      = TRUE,
  cache.lazy = FALSE
)
## Set dir relative to package
knitr::opts_knit$set(root.dir = '..')
```

# Intro to trees of cheem

## Requirements

```{R}
require("DALEX")
#require("treeshap")
require("spinifex")
require("tourr")
require("ggplot2")
## Especially, see basis_cheem() and view_cheem()
source("./apps/poc/trees_of_cheem.r")
if(F)
  file.edit("./apps/poc/trees_of_cheem.r")
## New spinifex ggproto_* api
source("./apps/poc/spinifex_ggproto.r")
if(F)
  file.edit("./apps/poc/spinifex_ggproto.r")
```

## Data setup

```{R}
str(tourr::flea) ## 74 x 6 numeric + 3 level class
dat <- spinifex::scale_sd(tourr::flea[, 1:6])
clas <- tourr::flea$species
new_obs <- 20
tgt_var <- clas == clas[new_obs] ## Or regression on a continuous var not in data.
```

## Find a start basis

```{r}
bas_cheem <- basis_cheem(
  data = dat,
  holdout_rownum = new_obs,
  target_var = tgt_var,
  class = clas,
  parts_type = "shap",
  basis_type = NULL)
print(bas_cheem) ## 1 dim
```

### DALEX::plot.predict_parts() symmetrically robust?

In the following evaluated R chunk (see code) we permute both the output DALEX::predict_parts columns and that of the input data. Both do not change the print order (though permuting the parts order does destroy the local attribution CI.)

```{r, eval=FALSE}
.parts <- attributes(bas_cheem)$predict_parts
plot(.parts)

## Explore the variable order assyemtry; does var order matter?
colnames(dat) ## 1) plot.predict_parts() doesn't print in variable nor ABC order.
set.seed(1010)
.col_shuffle <- sample(1:6, 6)
# .parts2 <- .parts[.col_shuffle, ] ## 2) shuffling a plot.predict_parts() obj doesn't change the print order (but has destroys the CI of the values.)
print("Shuffling the predict parts obj, obscurses the CI of the local attribution of the shap values, but not the order displayed. what if we shuffle the variables right out of the gate?")

## Setup dat2, with shuffled columns:
dat2 <- spinifex::scale_sd(tourr::flea[, .col_shuffle])
bas_cheem2 <- basis_cheem(
  data = dat2,
  holdout_rownum = new_obs,
  target_var = tgt_var,
  class = clas,
  parts_type = "shap",
  basis_type = NULL)
.parts2 <- attributes(bas_cheem2)$predict_parts

require("patchwork")
plot(.parts) + plot(.parts2)
print("Great! robust to both the varaible order both before and after")
```


## View a cheem basis

```{r}
plot(cheem_basis = bas_cheem, show_parts = TRUE)
# Error in names(x) <- value : 
#   'names' attribute [3] must be the same length as the vector [2]

#debugonce(plot.cheem_basis)
#debugonce(ggplot_tour) ## Error in if (n < 2) { : missing value where TRUE/FALSE needed
debugonce(array2df) ## Error: in 1L:n_frames : NA/NaN argument)
```


## Tour the space

### Short cheem tour, 4 target bases

```{r}
if(interactive())
  ?spinifex::play_tour_path()
## tour_path wants an array of target bases to tourr::interpolate() between.

## Let's naively find different cheem bases where SHAP is always on X and we find different first components for Y.
.bas_types <- c('RF_importance', 'olda', 'pca', 'odp', 'onpp')
tgt_bases <- array(NA, dim = c(dim(bas_cheem), length(.bas_types))) ## init
mute <- sapply(seq_along(.bas_types), function(i){
  tgt_bases[,, i] <<-
    bas_cheem <- basis_cheem(
      data = dat,
      class = clas,
      holdout_rownum = new_obs,
      target_var = tgt_var,
      ## X axis is always SHAPely values
      parts_type = "shap",
      ## Y axis components, varying across basis_types
      basis_type = .bas_types[i])
})
## Format for tourr::interpolate() 
tgt_bases <- as_history_array(tgt_bases, dat)

## Aesthetic init, Note that this is reused for the next several chunks in the `identity_args`
.colorPal <- RColorBrewer::brewer.pal(8, "Dark2")
.col <- .colorPal[as.integer(clas)]
.col[new_obs] <- "red"
.shapePal <- 15:17
.shp <- .shapePal[as.integer(clas)]
.shp[new_obs] <- 8 ## shape 8 is *, 4 is x
.alp <- rep(0.7, nrow(dat))
.alp[new_obs] <- .9
.siz <- rep(2, nrow(dat))
.siz[new_obs] <- 6

## Tour animation
spinifex::play_tour_path(
  tour_path = tgt_bases,
  data = dat,
  ## position of the basis unit circle
  ## expects:  "center", "left", "right", "bottomleft", "topright", or "off"
  axes = "left", 
  angle = .05, ## Step size distance [radians] between interpolation planes
  fps = 8, 
  ## list that would go inside ggplot2::geom_point(aes(X))
  aes_args = list(), 
  ## list that would go inside geom, outside aes: ggplot2::geom_point(aes(), X)
  identity_args = list(color = .col, shape = .shp,
                       size = .siz, alpha = .alp),
  ## list of items to "add" to a ggplot() + X
  ggproto = list(spinifex::theme_spinifex()),
  ## expects render_gganimate (.gif) or render_plotly (.html widget)
  render_type = render_plotly
)
## For saving, store output of spinifex::play_tour_path()
# ?gganimate::anim_save() ## with render_gganimate
# ?htmlwidgets::saveWidget() ## with render_gganimate
```

### Cheem tour, useing new api; `ggproto_*` 

Note that aesthetics are still a pain, for the oos obs case.

```{r}
## Sourced above:
#source("./apps/poc/spinifex_ggproto.r")
ggt <-
  ggplot_tour(tgt_bases, dat) +
  ggproto_basis_axes() +
  ggproto_data_points(identity_args = list(color = .col, shape = .shp,
                                           size = .siz, alpha = .alp)) +
  ggproto_data_background(gridlines = FALSE)

animate_plotly(ggt)
```

### Discrete view of cheem's target bases

```{r, fig.width=8}
## Sourced above:
#source("./apps/poc/spinifex_ggproto.r")
.parts <- attributes(bas_cheem)$predict_parts ## Note that array loses predicted parts
gg_ls <- list(plot(.parts))
mute <- sapply(seq_along(tgt_bases), function(i){
  .this_tgt_base <- tgt_bases[,, i]
  gg_ls[[i + 1]] <<-
    ggplot_tour(.this_tgt_base, dat) +
    ggproto_basis_axes() +
    ggproto_data_points(identity_args = list(color = .col, shape = .shp,
                                             size = .siz, alpha = .alp)) +
    ggproto_data_background(gridlines = FALSE)
})

.t <- theme(axis.title = element_text())
gg_ls[[1]]
gg_ls[[2]] + .t + labs(x = "shap", y = .bas_types[2 - 1])
gg_ls[[3]] + .t + labs(x = "shap", y = .bas_types[3 - 1])
gg_ls[[4]] + .t + labs(x = "shap", y = .bas_types[4 - 1])
gg_ls[[5]] + .t + labs(x = "shap", y = .bas_types[5 - 1])
gg_ls[[6]] + .t + labs(x = "shap", y = .bas_types[6 - 1])
``` 

### Frozen tour route

```{r, eval=FALSE}
if(interactive())
  ?tourr::frozen_tour

## Create a freezing matrix of dim == basis
freezer_mat <- matrix(NA, nrow = nrow(cheem_bas), ncol = ncol(cheem_bas))
## Try to freeze the first columns, the SHAP values.
freezer_mat[, 1] <- .99 * (1 / nrow(cheem_bas))

## Tour attempt erroring
if(interactive())
  tourr::animate_xy(data = dat,
                    tour_path = tourr::frozen_tour(frozen = freezer_mat),
                    start = cheem_bas)
## R> Error in svd(t(Fa) %*% Fz) : infinite or missing values in 'x'
```

